---
title: "Toronto Crime Analysis in R"
output: html_notebook
---

Data was obtained from: https://data.torontopolice.on.ca/datasets/mci-2014-to-2019/

### Date Preparation 🔬
```{r}
install.packages("tidyverse")
install.packages("data.table")
install.packages("lubridate")
install.packages("ggmaps")
install.packages("maps")
install.packages("mapdata")
install.packages("ggrepel")
```

```{r}
library(tidyverse) 
library(data.table) 
library(lubridate) 
library(ggmap) 
library(maps) 
library(mapdata)
library(ggrepel)
library(forcats)
```

```{r}
#Import data
crime <- as.data.frame(read.csv("toronto_crimes.csv"))
```

### Time to clean and select relevant variables 🧼

This dataset includes all the major crime indicators from 2014 - 2019 in Toronto, ON. 
```{r}
crime_selected <- select(crime, 'occurrencedate', 'occurrencemonth', 'occurrenceday', 'occurrencedayofweek', 'occurrenceyear', 'occurrencehour', 'offence', 'MCI', 'Neighbourhood', 'Long', 'Lat')
```

```{r}
#Remove the trailing zeros in the date column so we can filter the data by date
crime_selected$occurrencedate <- str_remove(crime_selected$occurrencedate, "00:00:00\\+00")

```
### Crime in 2017 🚓

How can we find the most occurring crimes within a certain year? 
```{r}
#Filter data between two occurrence dates. 
crime_2017 <- filter(crime_selected, `occurrencedate` >= as_date("2017-01-01"), `occurrencedate` <= as_date("2017-12-30"))

```
```{r}
#grouping by offence 
g2017 <- crime_2017 %>%group_by(offence) %>%
  summarise(total = n()) %>%
  distinct() %>%
  top_n(10)
```
```{r}
g2017 %>% 
  mutate(offence = fct_reorder(offence, total)) %>%
  ggplot( aes(x=offence, y=total)) +
    geom_bar(stat="identity", fill="#f68060", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") +
    labs(title="Top 10 Crimes Commited in 2017", x ="Offence", y = "Total Incidents") +
    theme_bw()
```
We can see above that assault is the most commited crime, followed by breaking and entering and theft of motor vehichle. 

### Top 5 Neighbourhoods With Highest Assault Incidents (2017)

What if we wanted to know which neighbour hoods had the most assault? 
```{r}
#Filter by date and offence
assault_total <- filter(crime_selected, `occurrencedate` >= as_date("2017-01-01"), `occurrencedate` <= as_date("2017-12-30"), offence == 'Assault')
```
```{r}
#group by neighbourhood
assaultTest <- assault_total  %>%
  group_by(occurrenceyear, Neighbourhood) %>%
  summarise(total = n()) %>%
  distinct() %>%
  top_n(5)
```
```{r}
assaultTest %>% 
  mutate(Neighbourhood = fct_reorder(Neighbourhood, total)) %>%
  ggplot(aes(x =Neighbourhood, y = total)) +
  geom_bar(stat="identity", fill="#93b793", alpha=.6, width=.4) +
    coord_flip() +
    xlab("") +
    labs(title="Top 5 Neighbourhoods With Highest Assault Incidents", x ="Neighbourhood", y = "Total Incidents") +
    theme_bw() +
  geom_label_repel(aes(label = total), size = 2.5)

```
From the plot above, it can be seen that the Waterfront Communities has the most assault incidents.


### Mapping Assault in Downtown Toronto 

```{r}

register_google(key = "")
```

```{r}
#gets the map of toronto
toronto <- get_map(location ="toronto", color = "bw",  zoom = 15)

#filter our data for only assault incidents in 2017 
mapping <- crime_2017 %>%
  select(offence, Long, Lat, occurrencedayofweek) %>%
  filter(offence == 'Assault' ) %>%
  na.omit()

```
```{r}
torontomap <- toronto %>%
  ggmap(base_layer =ggplot (aes(x = Long, y = Lat), data = mapping)) + 
  stat_density2d(aes(x = Long, y = Lat, fill = ..level.., alpha = ..level..),
  bins = 12, geom = "polygon", data = mapping ) +
  facet_wrap(~occurrencedayofweek, nrow = 2)  + scale_fill_gradient(low = "green", high = "red", 
    guide = FALSE) + scale_alpha(range = c(0, 0.3), guide = FALSE)

torontomap
```
From the above plots we can see that most of the assault incidents occur near Yonge & Dundas which has an active nightlife. 
We can see this pattern throughout the week because thats where the crimes are concentrated.  

